---
title: "GE_TimeSeries_Analysis"
author: "PraveenJoshi"
date: "August 9, 2017"
output: html_document
---

### Problem Description
 
 Build time series models in using GE stock data and the goal is to predict the price of a stock for future time period like 1 weeks ahead using time series algorithms and evaluating each of the
algorithms.


###Clear the Global Environment
```{r}
rm(list=ls(all=TRUE))
```
### Library Call
```{r}
library(forecast)
library(lubridate)
library(DataCombine)
library(imputeTS)
library(dplyr)
library(TTR)
library(graphics)
library(data.table)
library(plyr)
library(zoo)
```
### Read Data from CSV
```{r}
setwd("C:\\Users\\prave\\Desktop\\time-series-assignment-praveenjoshi01-master")
data = read.csv("BSE-BOM532309.csv")
```
### Observe the data
* Find how many unique attributes are there
* What are all the columns given in the data
```{r}
names(data)
##Summary
summary(data$Condition)
## Dimensions
dim(data)
```
## Basic data View
```{r}
head(data,30)
```
## Summary of the entire data set
```{r}
summary(data)
str(data)
```
### Observation & Analysis 
* We observe that there is different opening price for the GE stock at different day.
* R has read the data column either as factor we need convert it to date format
```{r}
data$Date=as.Date(data$Date,format="%Y-%m-%d")
```
### Missing values in time series
* Observe the data to find if any
```{r}
sum(is.na(data))
colSums(is.na(data))
str(data$Date)
```

### Creating a data frame for sequence of Days 
```{r }
seq <- data.frame("observedDays"=seq(1,4159))
```
#### Merge the data frames of seq of dates with the dates of stocks obtained from BSE
```{r }
RtData2=cbind(seq,data)
head(RtData2)
```

### Extracting out relative dataSet for further Analysis.
```{r}
#zoo package
RtData2<- RtData2[,c(1,3)]
head(RtData2)
```


### Splitting of the Data
* Random split is not possible because here we will be in need of sequence where by we miss the data points
* splitting is done by sequential splitting
```{r}
Train=RtData2[10:4159,]
Test=RtData2[1:9,]
```

### converting into time series 
* our target variable is opening price
```{r}
StockSeries <- ts(Train$Open,frequency =5)
```

### Vizualize the time series Data
```{r}
plot(StockSeries,
     type="l",
     lwd=3,
     col="red",
     xlab="yearly",
     ylab="Price",
     main="Time series plot for Book-xyzabc")
```
### Decomposed Time Series
* Decompose will provide us with the info on seasonality,trend and randomness
```{r}
StockSeriesdecomposed=decompose(StockSeries)
plot(StockSeriesdecomposed,col="Red")
```
### ACF,PACF 
* Autocorrelation is the linear dependence of a variable with itself at two points in time
* For stationary processes, autocorrelation between any two observations only depends on the time lag h between them
*  Partial autocorrelation is the autocorrelation between yt and yt–h after removing any linear dependence on y1,y2, ..., yt–h+1
```{r}
par(mfrow=c(2,2))
acf(StockSeries,lag=30)
pacf(StockSeries,lag=30)
```


```{r }
StockSeries1 <- ts(Train$Open, frequency =1)
plot(StockSeries1,xlab = "Weekly",ylab="Price")
acf(StockSeries1,lag=30)
pacf(StockSeries1,lag=30)
```

### Stationarize by differencing

```{r}
par(mfrow=c(2,3))
plot(diff(StockSeries1,lag = 1),type="l"); acf(diff(StockSeries1,lag = 1),lag=30) ;pacf(diff(StockSeries1,lag = 1),lag=30)
plot(diff(StockSeries1,lag=2),type="l");  acf(diff(StockSeries1,lag = 2),lag=30); pacf(diff(StockSeries1,lag = 2),lag=30)
```
* one lag has stationarize the data we can use ndiffs of forecast package to check no of differences required to      stationarize the data
```{r}
ndiffs(StockSeries1)
```


### Modelling  the time series using simple moving averages
* Time series Price has trend 
* Modelling the time series behaviour by simple moving averages
```{r}
fitsma <- SMA(StockSeries,n=2)
pred<-forecast(fitsma,h=4)
plot(pred)
lines(fitsma,col="red")
```
### Define the metric MAPE 
```{r}
smaMape <- mean(abs((StockSeries[2:length(StockSeries)]-fitsma[2:length(StockSeries)])/StockSeries[2:length(StockSeries)]))
smaMape
```

### Weighted Moving Averages
```{r}
fitwma<- WMA(StockSeries,n=2,1:2)
wmaMape <- mean(abs((StockSeries[2:length(StockSeries)]-fitwma[2:length(StockSeries)])/StockSeries[2:length(StockSeries)]))
wmaMape
pred<-forecast(fitwma,h=4)
plot(pred)
lines(fitwma,col="yellow")
```

### Exponential Moving Averages
```{r}
fitEma <- EMA(StockSeries, n = 2)
emaMape <- mean(abs((StockSeries[2:length(StockSeries)]-fitEma[2:length(StockSeries)])/StockSeries[2:length(StockSeries)]))
emaMape
pred<-forecast(fitEma,h=4)
plot(pred)
lines(fitEma,col="purple")
```

## Build a HoltWinters model  with trend 
```{r}
holtpriceforecast <- HoltWinters(Train$Open,gamma=FALSE)
head(holtpriceforecast$fitted)
```
## HoltWinters model  with trend  and Seasonality

```{r}
priceholtforecast <-
HoltWinters(StockSeries, beta=TRUE, gamma=TRUE, seasonal="additive")
head(priceholtforecast$fitted)
```


### Prediction on the Train
```{r}
holtforecastTrain <- data.frame(priceholtforecast$fitted)
holtforecastTrainpredictions <- holtforecastTrain$xhat
head(holtforecastTrainpredictions)
```

### Prediction on test data
```{r}
priceforecast<-forecast(priceholtforecast,h = 8)
plot(priceforecast)

```


### Arima Models
```{r}

model1 <- arima(StockSeries,c(0,0,0))
model1
acf(StockSeries) 
pacf(StockSeries)
plot(StockSeries)
#non seasonal differentiating
ndiffs(StockSeries)
#seasonal differentiating
nsdiffs(StockSeries)
```
## Considering the difference from the graph as d=1 to stationarize
```{r}
model2 <- arima(StockSeries,c(0,1,0))
model2
acf(diff(StockSeries,lag = 1))
pacf(diff(StockSeries,lag = 1))
plot(diff(StockSeries))
```
* plot has still non stationary behaviour another difference can stationarize it 
```{r}
model3 <- arima(StockSeries,c(0,2,0))
model3
plot(diff(StockSeries,differences = 2))
acf(diff(StockSeries,differences = 2))
pacf(diff(StockSeries,differences = 2))
```
* Observing the acf and pacf there is significant lag in acf and also in pacf that has to be taken care 
```{r}
model4 <- arima(StockSeries,c(1,1,1))
model4


```
## Plots of the models

```{r}
par(mfrow=c(2,2))
plot(model1$residuals,ylim=c(-50,50))
plot(model2$residuals,ylim=c(-50,50))
plot(model3$residuals,ylim=c(-50,50))
plot(model4$residuals,ylim=c(-50,50))
```
###  Auto Arima
```{r}
MODEL_ARIMA <- auto.arima(StockSeries, ic='aic')
summary(MODEL_ARIMA)
```

### Forecast on the models 
```{r}
pricearimaforecasts1 <- forecast(model1, h=4)
plot(pricearimaforecasts1)
pricearimaforecast3 <- forecast(model3, h=4)
plot(pricearimaforecast3)
pricearimaforecasts_autArima<- forecast(MODEL_ARIMA,h=4)
plot(pricearimaforecasts_autArima,flwd = 2)
summary(MODEL_ARIMA)

```

*Comparing the AIC Auto Arima was giving good results for the product item considered